### 2014-10-01 tex/proofs.tex
Rewrote theorem 1

\subsection{Theorems}
\begin{lemma}[Location Mutual Exclusion in $\rightarrow^*_S$]
Given a location mutual exclusive heap, $(\cfgnt{L}_0\ \cfgnt{R}_0)$, from a
summary state with a reference $\cfgnt{r}$ and field $\cfgnt{f}$, the
new heap, $(\cfgnt{L}^\prime\ \cfgnt{R}^\prime)$, from the summary
initialization machine, $(\cfgnt{L}_0\ \cfgnt{R}_0\ \cfgnt{r}\ \cfgnt{f})
\rightarrow^*_S
(\cfgnt{L}^\prime\ \cfgnt{R}^\prime\ \cfgnt{r}\ \cfgnt{f})$, after any number of machine steps, is also
location mutual exclusive.
\end{lemma}
\begin{proof}
Induction over the number of machine steps on the summary initialization machine in \figref{fig:symInit}.

\noindent\textbf{Base Case}. The machine takes one step: $(\cfgnt{L}_0\ \cfgnt{R}_0\ \cfgnt{r}\ \cfgnt{f})
\rightarrow_S
(\cfgnt{L}_1 \cfgnt{R}_1\ \cfgnt{r}\ \cfgnt{f})$. Let $\Lambda = \mathbb{UN}(\cfgnt{L}_0, \cfgnt{R}_0, \cfgnt{r},
\cfgnt{f})$ be the set of uninitialized locations. If $\Lambda = \emptyset$, then the \texttt{Summarize (end)}
rule is active and $(\cfgnt{L}_1\ \cfgnt{R}_1) = (\cfgnt{L}_0\ \cfgnt{R}_0)$, which is location mutual exclusive by definition.

If $\Lambda \neq \emptyset$, then the \texttt{Summarize} rule is
active, and each new constraint location pair must be considered
individually. These pairs are partitioned into the sets
$\theta_\mathit{null}$, $\theta_\mathit{new}$,
$\theta_\mathit{alias}$, and $\theta_\mathit{orig}$ by the rule. 
\begin{itemize}
\item Any member of $\theta_\mathit{null}$, $\theta_\mathit{new}$, and
  $\theta_\mathit{alias}$ has a constraint of the form $\phi_x \wedge
  \ldots$ while any member of $\theta_\mathrm{orig}$ has a constraint
  of the form $\neg \phi_x \wedge \ldots$. As every
  $(\phi_\mathit{orig}\ \cfgnt{l}_\mathit{orig}) \in
  \cfgnt{L}_n(\cfgnt{R}_n(\cfgnt{l}_x,\cfgnt{f}))$ is location mutual exclusive
  by definition, members of $\theta_\mathit{orig}$ are location
  mutual exclusive with each other and with members in $\theta_\mathit{null}$,
  $\theta_\mathit{new}$, and $\theta_\mathit{alias}$.
\item Any member of $\theta_\mathit{null}$ has the form $\ldots \wedge
  \cfgnt{r}_f = \cfgnt{r}_\mathit{null}$ while any member of
  $\theta_\mathit{new}$ and $\theta_\mathit{alias}$ has the form
  $\ldots \wedge \cfgnt{r}_f \neq \cfgnt{r}_\mathit{null} \wedge
  \ldots$. As $\theta_\mathit{null}$ is a singleton set, its location is
  location mutual exclusive from those in $\theta_\mathit{new}$ and
  $\theta_\mathit{alias}$.
\item Any member of $\theta_\mathit{new}$ has a constraint of the form
  $\ldots \wedge ( \wedge_{(
  \cfgnt{r}_a,\ \phi_a,\ l_a) \in \rho}
  \cfgnt{r}_f \ne \cfgnt{r}_a) )$ to assert it does not
  alias anything while any member of $\theta_\mathit{alias}$ has
  the form $\ldots \wedge \cfgnt{r}_f = \cfgnt{r}_a \wedge \ldots$ for
  some alias $\cfgnt{r}_a$ with both partitions reasoning over the same set of aliases $\rho$.  As $\theta_\mathit{new}$ is a singleton set, its
  location is location mutual exclusive from any member of
  $\theta_\mathit{alias}$.
\item Any member of $\theta_\mathit{alias}$ has the form
\[\ldots \wedge \cfgnt{r}_f = \cfgnt{r}_a \wedge ( \wedge_{( \cfgnt{r}^{\prime}_a\ \phi^{\prime}_a\ l^{\prime}_a)  \in \rho\ ( \cfgnt{r}^\prime_a \neq \cfgnt{r}_a) } \cfgnt{r}_f \neq \cfgnt{r}^{\prime}_a)\]
Each is mutually exclusive over the particular alias it asserts.
\end{itemize}
Every member in $\theta = \theta_\mathit{null} \cup
\theta_\mathit{new} \cup \theta_\mathit{alias} \cup
\theta_\mathit{orig}$ is thus location mutual exclusive making $(\cfgnt{L}_1 \cfgnt{R}_1\ \cfgnt{r}\ \cfgnt{f})$ location mutual exclusive as well.

\noindent\textbf{Inductive Step}. The machine takes $n$-steps:
\[
(\cfgnt{L}_0\ \cfgnt{R}_0\ \cfgnt{r}\ \cfgnt{f}) \rightarrow_S (\cfgnt{L}_1\ \cfgnt{R}_1\ \cfgnt{r}\ \cfgnt{f}) \rightarrow_S \ldots \rightarrow_S
(\cfgnt{L}_n\ \cfgnt{R}_n\ \cfgnt{r}\ \cfgnt{f})
\]
By the induction hypothesis, $(\cfgnt{L}_n\ \cfgnt{R}_n)$ is location
mutual exclusive. This matches the base case, in that the heap on the
left side of the transition is location mutual exclusive, and by the
same argument as in the base case,
$(\cfgnt{L}_{n+1}\ \cfgnt{R}_{n+1})$ is thus location mutual exclusive.
\begin{comment}
The Field Access rule uses the Summarize rule to describe the relation between $\cfgnt{L}_{s}$ and $\cfgnt{L}_{s^\prime}$. Because the summarize rule is essentially a fixed-point computation, we can reason about it as a machine that produces a sequence of intermediate states $s_1, s_2, ... , s_n$. We will use an inductive argument to show that the mutual exclusion property holds for any of those intermediate states. First, let state $s_\alpha$ be any state for which the property in Theorem~\ref{thm:mutex} holds, and let $s_\beta$ be the intermediate state where $s_\alpha \rightarrow_S s_\beta$. If all fields are initialized, then $s_\alpha$ and $s_\beta$ are identical, so the mutually exclusive property holds for $s_\beta$. 
Otherwise, there must be some location $\cfgnt{l}_{un} \in \cfgnt{L}_\alpha(r)$ for which field f contains a reference $r_a$ that could feasibly point to the uninitialized location. In this case, a new reference is created and a single entry is added to $\cfgnt{L}_\beta$. The elements of the set $\theta_{all} = \cfgnt{L}_\beta(\cfgnt{R}_\beta(\cfgnt{l}_{un},\cfgnt{f}))$ can be divided into four distinct subsets, $\theta_{cur}$, $\theta_{null}$, $\theta_{new}$, and $\theta_{alias}$. The set $\theta_{cur}$ contains the pairs containing locations from the reference $r_a$, $\theta_{null}$ contains the pair with the null location, $\theta_{new}$ contains a pair with a new location, and $\theta_{alias}$ contains pairs representing alias locations. Any two constraints in $\theta_{cur}$ are guaranteed to be pairwise mutually exclusive because of the inductive hypothesis, and any single constraint is guaranteed to be pairwise mutually exclusive with any constraint in $\theta_{null}$, $\theta_{new}$, or $\theta_{alias}$ because constraints in $\theta_{cur}$ contain the conjunct $\neg \phi$, while constraints in the other three sets are conjoined with $\phi$. The set $\theta_{null}$ contains only one element, which in turn contains the constraint $r_f = r_{null}$, which is guaranteed to be mutually exclusive with any constraint in $\theta_{new}$, or $\theta_{alias}$ because constraints in those sets contain the conjunct $r_f \neq r_{null}$. The set $\theta_{new}$ contains one element, which in turn contains the constraint containing the conjunct $ \wedge_{\lp \cfgnt{r}^\prime_a,\ \phi^\prime_a,\ l^\prime_a\rp  \in \rho} \cfgnt{r}_f \ne \cfgnt{r}^\prime_a$, which is guaranteed to be mutually exclusive with any constraint in $\theta_{alias}$, because each constraint in that set contains the conjunct $r_f = r_a$ for some $(r_a\ \phi_a\ l_a)\in \rho$. Any two elements of the set $\theta_{alias}$ are guaranteed to pairwise mutually exclusive because for each constraint that contains the conjunct $r_f = r_a$, every other constraint contains the conjunct $r_f \ne r_a$. Thus, all constraints from all pairs in $\theta{all}$ are guaranteed to be mutually exclusive, and the property from Theorem~\ref{thm:mutex} holds for intermediate state $s_\beta$.

@@

\begin{theorem}[Mutual Exclusion]
\label{thm:mutex}
If we have a reachable summary state $s_s = ( \cfgnt{L}_s\ \cfgnt{R}_s\ \phi_s\ \eta_s\ \cfgnt{e}_s\ \cfgnt{k}_s )$, then for any reference $r \in \cfgnt{L}_s^\leftarrow$
$$ 
\begin{array}{l}
\forall (\phi_\alpha\ \cfgnt{l}_\alpha) \in \cfgnt{L}_s(\cfgnt{r})\ (\forall (\phi_\beta\ \cfgnt{l}_\beta) \in \cfgnt{L}_s(\cfgnt{r})\ ( \\
\ \ \ \ (\cfgnt{l}_\alpha \neq \cfgnt{l}_\beta) \Rightarrow ((\phi_\alpha \wedge \phi_\beta) = \cfgt{false})))
\end{array}
$$
\end{theorem}
\begin{proof}
The proof will proceed inductively.

>>>>

\subsection{Theorems}
\begin{lemma}[$\rightarrow^*_S$ preserves location mutual exclusion]
Given a location mutual exclusive heap, $(\cfgnt{L}_0\ \cfgnt{R}_0)$, from a
summary state with a reference $\cfgnt{r}$ and field $\cfgnt{f}$, the
new heap, $(\cfgnt{L}^\prime\ \cfgnt{R}^\prime)$, from the summary
initialization machine, $(\cfgnt{L}_0\ \cfgnt{R}_0\ \cfgnt{r}\ \cfgnt{f})
\rightarrow^*_S
(\cfgnt{L}^\prime\ \cfgnt{R}^\prime\ \cfgnt{r}\ \cfgnt{f})$, after any number of machine steps, is also
location mutual exclusive.
\end{lemma}
\begin{proof}
Induction over the number of machine steps on the summary initialization machine in \figref{fig:symInit}.

\noindent\textbf{Base Case}. The machine takes one step: $(\cfgnt{L}_0\ \cfgnt{R}_0\ \cfgnt{r}\ \cfgnt{f})
\rightarrow_S
(\cfgnt{L}_1\ \cfgnt{R}_1\ \cfgnt{r}\ \cfgnt{f})$. Let $\Lambda = \mathbb{UN}(\cfgnt{L}_0, \cfgnt{R}_0, \cfgnt{r},
\cfgnt{f})$ be the set of uninitialized locations. If $\Lambda = \emptyset$, then the \texttt{Summarize (end)}
rule is active and $(\cfgnt{L}_1\ \cfgnt{R}_1) = (\cfgnt{L}_0\ \cfgnt{R}_0)$, which is location mutual exclusive by the initial conditions in the lemma.

If $\Lambda \neq \emptyset$, then the \texttt{Summarize} rule is
active, and each new constraint location pair must be considered
individually. These pairs are partitioned into the sets
$\theta_\mathit{null}$, $\theta_\mathit{new}$,
$\theta_\mathit{alias}$, and $\theta_\mathit{orig}$ by the rule. 
\begin{itemize}
\item Any member of $\theta_\mathit{null}$, $\theta_\mathit{new}$, and
  $\theta_\mathit{alias}$ has a constraint of the form $\phi_x \wedge
  \ldots$ while any member of $\theta_\mathrm{orig}$ has a constraint
  of the form $\neg \phi_x \wedge \ldots$. As every
  $(\phi_\mathit{orig}\ \cfgnt{l}_\mathit{orig}) \in
  \cfgnt{L}_n(\cfgnt{R}_n(\cfgnt{l}_x,\cfgnt{f}))$ is location mutual exclusive
  by the initial conditions on $(\cfgnt{L}_0\ \cfgnt{R}_0)$ in the lemma statement, members of $\theta_\mathit{orig}$ are location
  mutual exclusive with each other and with members in $\theta_\mathit{null}$,
  $\theta_\mathit{new}$, and $\theta_\mathit{alias}$.
\item Any member of $\theta_\mathit{null}$ has the form $\ldots \wedge
  \cfgnt{r}_f = \cfgnt{r}_\mathit{null}$ while any member of
  $\theta_\mathit{new}$ and $\theta_\mathit{alias}$ has the form
  $\ldots \wedge \cfgnt{r}_f \neq \cfgnt{r}_\mathit{null} \wedge
  \ldots$. As $\theta_\mathit{null}$ is a singleton set, its location is
  location mutual exclusive from those in $\theta_\mathit{new}$ and
  $\theta_\mathit{alias}$.
\item Any member of $\theta_\mathit{new}$ has a constraint of the form
  $\ldots \wedge ( \wedge_{(
  \cfgnt{r}_a,\ \phi_a,\ l_a) \in \rho}
  \cfgnt{r}_f \ne \cfgnt{r}_a) )$ to assert it does not
  alias anything while any member of $\theta_\mathit{alias}$ has
  the form $\ldots \wedge \cfgnt{r}_f = \cfgnt{r}_a \wedge \ldots$ for
  some alias $\cfgnt{r}_a$ with both partitions reasoning over the same set of aliases $\rho$.  As $\theta_\mathit{new}$ is a singleton set, its
  location is location mutual exclusive from any member of
  $\theta_\mathit{alias}$.
\item Any member of $\theta_\mathit{alias}$ has the form
\[\ldots \wedge \cfgnt{r}_f = \cfgnt{r}_a \wedge ( \wedge_{( \cfgnt{r}^{\prime}_a\ \phi^{\prime}_a\ l^{\prime}_a)  \in \rho\ ( \cfgnt{r}^\prime_a \neq \cfgnt{r}_a) } \cfgnt{r}_f \neq \cfgnt{r}^{\prime}_a)\]
Each is mutually exclusive over the particular alias it asserts.
\end{itemize}
Every member in $\theta = \theta_\mathit{null} \cup
\theta_\mathit{new} \cup \theta_\mathit{alias} \cup
\theta_\mathit{orig}$ is thus location mutual exclusive making $(\cfgnt{L}_1\ \cfgnt{R}_1\ \cfgnt{r}\ \cfgnt{f})$ location mutual exclusive as well.

\noindent\textbf{Inductive Step}. The machine takes $n$-steps:
\[
(\cfgnt{L}_0\ \cfgnt{R}_0\ \cfgnt{r}\ \cfgnt{f}) \rightarrow_S (\cfgnt{L}_1\ \cfgnt{R}_1\ \cfgnt{r}\ \cfgnt{f}) \rightarrow_S \ldots \rightarrow_S
(\cfgnt{L}_n\ \cfgnt{R}_n\ \cfgnt{r}\ \cfgnt{f})
\]
By the induction hypothesis, $(\cfgnt{L}_n\ \cfgnt{R}_n)$ is location
mutual exclusive. This matches the base case, in that the heap on the
left side of $\rightarrow_S$ is location mutual exclusive, and by the
same argument as in the base case,
$(\cfgnt{L}_{n+1}\ \cfgnt{R}_{n+1})$ is thus location mutual exclusive.

\begin{comment}
% Initial proof text taken from the theorem
The Field Access rule uses the Summarize rule to describe the relation between $\cfgnt{L}_{s}$ and $\cfgnt{L}_{s^\prime}$. Because the summarize rule is essentially a fixed-point computation, we can reason about it as a machine that produces a sequence of intermediate states $s_1, s_2, ... , s_n$. We will use an inductive argument to show that the mutual exclusion property holds for any of those intermediate states. First, let state $s_\alpha$ be any state for which the property in Theorem~\ref{thm:mutex} holds, and let $s_\beta$ be the intermediate state where $s_\alpha \rightarrow_S s_\beta$. If all fields are initialized, then $s_\alpha$ and $s_\beta$ are identical, so the mutually exclusive property holds for $s_\beta$. 
Otherwise, there must be some location $\cfgnt{l}_{un} \in \cfgnt{L}_\alpha(r)$ for which field f contains a reference $r_a$ that could feasibly point to the uninitialized location. In this case, a new reference is created and a single entry is added to $\cfgnt{L}_\beta$. The elements of the set $\theta_{all} = \cfgnt{L}_\beta(\cfgnt{R}_\beta(\cfgnt{l}_{un},\cfgnt{f}))$ can be divided into four distinct subsets, $\theta_{cur}$, $\theta_{null}$, $\theta_{new}$, and $\theta_{alias}$. The set $\theta_{cur}$ contains the pairs containing locations from the reference $r_a$, $\theta_{null}$ contains the pair with the null location, $\theta_{new}$ contains a pair with a new location, and $\theta_{alias}$ contains pairs representing alias locations. Any two constraints in $\theta_{cur}$ are guaranteed to be pairwise mutually exclusive because of the inductive hypothesis, and any single constraint is guaranteed to be pairwise mutually exclusive with any constraint in $\theta_{null}$, $\theta_{new}$, or $\theta_{alias}$ because constraints in $\theta_{cur}$ contain the conjunct $\neg \phi$, while constraints in the other three sets are conjoined with $\phi$. The set $\theta_{null}$ contains only one element, which in turn contains the constraint $r_f = r_{null}$, which is guaranteed to be mutually exclusive with any constraint in $\theta_{new}$, or $\theta_{alias}$ because constraints in those sets contain the conjunct $r_f \neq r_{null}$. The set $\theta_{new}$ contains one element, which in turn contains the constraint containing the conjunct $ \wedge_{\lp \cfgnt{r}^\prime_a,\ \phi^\prime_a,\ l^\prime_a\rp  \in \rho} \cfgnt{r}_f \ne \cfgnt{r}^\prime_a$, which is guaranteed to be mutually exclusive with any constraint in $\theta_{alias}$, because each constraint in that set contains the conjunct $r_f = r_a$ for some $(r_a\ \phi_a\ l_a)\in \rho$. Any two elements of the set $\theta_{alias}$ are guaranteed to pairwise mutually exclusive because for each constraint that contains the conjunct $r_f = r_a$, every other constraint contains the conjunct $r_f \ne r_a$. Thus, all constraints from all pairs in $\theta{all}$ are guaranteed to be mutually exclusive, and the property from Theorem~\ref{thm:mutex} holds for intermediate state $s_\beta$.

@@

\begin{theorem}[$\rightarrow_\varsigma$ preserves location mutual exclusion]
\label{thm:mutex}
Given a summary state, $s_s = ( \cfgnt{L}_s\ \cfgnt{R}_s\ \phi_s\ \eta_s\ \cfgnt{e}_s\ \cfgnt{k}_s )$, such that $s_s \rightarrow_\varsigma s_s^\prime$, where $s_s^\prime = ( \cfgnt{L}_s^\prime\ \cfgnt{R}_s^\prime\ \phi_s^\prime\ \eta_s^\prime\ \cfgnt{e}_s^\prime\ \cfgnt{k}_s^\prime )$, if the heap $(\cfgnt{L}_s\ \cfgnt{R}_s)$ is location mutual exclusive then the heap after the transition, $(\cfgnt{L}_s^\prime\ \cfgnt{R}_s^\prime)$ is also location mutual exclusive.
\begin{comment}
If we have a reachable summary state $s_s = ( \cfgnt{L}_s\ \cfgnt{R}_s\ \phi_s\ \eta_s\ \cfgnt{e}_s\ \cfgnt{k}_s )$, then for any reference $r \in \cfgnt{L}_s^\leftarrow$
$$ 
\begin{array}{l}
\forall (\phi_\alpha\ \cfgnt{l}_\alpha) \in \cfgnt{L}_s(\cfgnt{r})\ (\forall (\phi_\beta\ \cfgnt{l}_\beta) \in \cfgnt{L}_s(\cfgnt{r})\ ( \\
\ \ \ \ (\cfgnt{l}_\alpha \neq \cfgnt{l}_\beta) \Rightarrow ((\phi_\alpha \wedge \phi_\beta) = \cfgt{false})))
\end{array}
$$
\end{comment}
\end{theorem}
\begin{proof}
The proof will proceed inductively.
